### 2022-04-27 acks/acks.tex
more
wan comments

After reading many preface or acknowledgments chapters I realized that there
is a discrepancy between what research shows candidate's experience of finishing
a PhD is and what is actually written in these parts of theses. An
article by Levecque et al. \cite{labhg-17-phdstudents} states that one in two
PhD students experiences psychological distress, while one in three is at risk
of a common psychiatric disorder. In the light of this fact, I want to write
this chapter not only to the people whose help and work indebted me, but also to
(future) PhD students that, after looking up a detail in this thesis,
voyeuristically read this chapter (like I did many times for other theses).

@@

The fact that I was surrounded by very successful postdocs and PhD students on
later years of their studies made me feel like I was the only one having this
problem. To the students reading this chapter, I would very much like to assure
you that you are not.

What helped me go out of the self-doubting loop was actually taking action. From
the simplest ones, like saying ``I don't understand'' out loud instead of
silently nodding when I have a trouble following scientific discussion, to the
ones with more impact on my career like actively looking for projects I could
join while I did not have much work on my plate. As the years went on I felt
that I could more openly talk about my insecurities and I found out that many
people in my situation have them as well.

What made my PhD journey more enjoyable are undoubtedly the people I worked
with. I want to thank the whole Theoretical Computer Science group at the VU
which was there with me from the early days of my master studies. They helped me
organize my master programme so that I follow more theory courses which most
likely lead  me to start an academic career. I also want to thank other postdocs
and PhDs that studied during my PhD or are still studying. Most notably, I would
like to thank Alexander Bentkamp, one of the most gifted mathematicians I know.
I coauthored many of my papers with him, each of which was a pleasure to work
on. I would also like to thank all of my collaborators and people that suggested
improvements to the articles I wrote.

I want to thank my daily supervisor, Jasmin Blanchette. He always says that the
main product of a PhD is not a thesis but a person you come out as after four
years. With his guidance, I learned to approach things more rigorously, slower,
and with more attention. He also helped me tremendously  with my writing and
presentation skills. Even though he tried very hard, he did not teach me to use
articles (i.e., English articles ``the'', ``a'', or ``an'') though, and
capitulated by adding them to many of the articles I wrote.

@@

Lastly, I would like to thank Martijn. He was besides me for the last four years
and helped me overcome many of the doubts and issues I mentioned. He also always
patiently waits until my medication or second glass of Grüner Veltliner kick in
(not at the same time). I also want to thank Jelisaveta and Tara, which
sometimes had to deal with me when medication or wine did not kick in. All my other friends,
in Serbia or the Netherlands, thank you for your support.

>>>>

After reading many preface or acknowledgments chapters I realized that there
is a discrepancy between what research shows candidate's experience of finishing
a PhD is and what is actually written in these parts of theses. An
article by Levecque et al. \cite{labhg-17-phdstudents} states that one in two
PhD students experiences psychological distress, while one in three is at risk
of a common psychiatric disorder. In light of this fact, I want to write
this chapter not only to the people whose help and work indebted me, but also to
(future) PhD students that, after looking up a detail in this thesis,
voyeuristically read this chapter (like I did many times for other theses).

@@

The fact that I was surrounded by very successful postdocs and PhD students in
later years of their studies made me feel like I was the only one having this
problem. To the students reading this chapter, I would very much like to assure
you that you are not.

What helped me get out of the self-doubting loop was actually taking action. From
the simplest ones, like saying ``I don't understand'' out loud instead of
silently nodding when I had trouble following scientific discussion, to the
ones with more impact on my career like actively looking for projects I could
join while I did not have much work on my plate. As the years went on I felt
that I could more openly talk about my insecurities and I found out that many
people in my situation have them as well.

What made my PhD journey more enjoyable are undoubtedly the people I worked
with. I want to thank the whole Theoretical Computer Science group at the VU
which was there with me from the early days of my master studies. They helped me
organize my master programme so that I follow more theory courses which most
likely led  me to start an academic career. I also want to thank other postdocs
and PhDs that studied during my PhD or are still studying. Most notably, I would
like to thank Alexander Bentkamp, one of the most gifted mathematicians I know.
I coauthored many of my papers with him, each of which was a pleasure to work
on. I would also like to thank all of my collaborators and people that suggested
improvements to the articles I wrote.

I want to thank my daily supervisor, Jasmin Blanchette. He always says that the
main product of a PhD is not a thesis but the person you come out as after four
years. With his guidance, I learned to approach things more rigorously, slower,
and with more attention. He also helped me tremendously  with my writing and
presentation skills. Even though he tried very hard, he did not teach me to use
articles (i.e., English articles ``the'', ``a'', or ``an'') though, and
capitulated by adding them to many of the articles I wrote.

@@

Predrag Jani\v{c}i\'{c}, my professor from University of Belgrade who 
inspired me to deepen my knowledge in various areas of computer science, suggested
many improvements to the text of this thesis. I thank him as well.

Lastly, I would like to thank Martijn. He was besides me for the last four years
and helped me overcome many of the doubts and issues I mentioned. He also always
patiently waits until my medication or second glass of Grüner Veltliner kicks in
(not at the same time). I also want to thank Jelisaveta and Tara, which
sometimes had to deal with me when medication or wine did not kick in. All my other friends,
in Serbia or the Netherlands, thank you for your support.

### 2022-04-27 conclusion/conclusion.tex
more
wan comments

\ourpara{Higher-Order Unification} The complete variant of unification procedure
described in Chapter \ref{ch:unif} is proved complete with substantial
restrictions on the order and kind of applied rules. To further
remove redundancy, we plan to find alternative, less explosive unification rules
that do not compromise completeness. Furthermore, we want to investigate if
there are more combinations of unification rules that lead to redundant unifiers
and find ways to remove those redundancy-inducing combinations.

We also described only one pragmatic variant of the unification procedure. This
variant was designed after we sifted through hundreds of benchmarks to get the
taste of what unifiers occur in practice. It is possible that this pragmatic variant
can be further tweaked to perform more efficiently without loosing many useful unifiers.

Lastly, we fixed the order in which Zipperposition applies different unification
rules. It would be interesting to see if applying imitation before projection
(or vice versa) yields performance improvements on some kinds of problems. 
We also postponed the application of the most explosive rules such as iteration
to the very end. However, it would not be surprising that there are many hard hand-crafted
problems on which applying these rules would lead to a proof more quickly.

\ourpara{Pragmatic Techniques and Heuristics } Many of the Boolean reasoning
techniques described in Chapter \ref{ch:bools} were either inspired by the ones
implemented in traditional higher-order theorem provers or by the unsolved TPTP
benchmarks. Since the work present in Chapter \ref{ch:bools} anticipated the
complete higher-order calculus that could make these techniques obsolete, we
implemented only the most successful ones that are easy to port to
saturation-based provers.  Most notably, unlike Satallax
\cite{cb-2013-satallax}, Zipperposition does not use the fact that there only
finitely many functions of a type built only using function type constructor
($\rightarrow$) and Boolean type ($o$). Implementing techniques for efficient
enumeration of all Boolean functions of a given type could help Zipperposition
prove problems that are still out of its reach, but can be solved by Satallax.

In Chapter \ref{ch:ho-techniques} we explored how we keep the explosion inherent
to \osup{} calculus under control. Due to time and resource
constraints, in the evaluation of our new given-clause loop we left some
parameters that control which streams are queried fixed. We plan to tune these
parameters in the future.  Similarly, we provided only some of the clause
priority functions that rely on higher-order features. We plan to investigate
more proofs in detail to create new priority functions that more cautiously
separate less and more useful higher-order inferences. Lastly, we observed that
use of Ehoh as a backend greatly improves the performance of Zipperposition. We
plan to further investigate the proofs that Ehoh returns and find out how
Zipperposition can find them without the use of a backend.

\ourpara{Further Extensions of \ehohii{}} \ehohii{} outperforms other provers on
problems coming proof assistants. However, its performance on TPTP problems is
slightly behind Zipperposition. We plan to further investigate the benchmarks on
which \ehohii{} fails, but are proved by Zipperposition. One of the first
features we plan implement is dynamic clausification, which can possibly help
improve \ehohii{}'s performance on first-order problems as well. We also plan to
experiment with implementing some of the \osup{} rules necessary for
completeness such as \infname{FluidSup}. Heuristics of \ehohii{} are trained on
a subset of Sledgehammer and TPTP problems. We plan to use other benchmark sets
such as GRUNGE to make \ehohii{} efficient on a wider range of higher-order
benchmarks.

\ourpara{SAT-inspired eliminations for superposition} We already implemented
some of the techniques described in Chapter~\ref{ch:satfol} in E. However, this
implementation is not properly tested and optimized. We plan to put finishing
touches to this implementation and test the effects of implemented preprocessing
techniques on E's performance. There are more techniques used in SAT solving,
notably bounded variable addition \cite{mhb-12-reencoding} and blocked clause
addition \cite{ok-99-er}, that we did not port to first-order logic. We want to
investigate if these techniques can be lifted to first-order logic. Lastly, we
already started work on porting existing techniques to higher-order logic.

>>>>

\ourpara{Higher-Order Unification} The complete variant of the unification procedure
described in Chapter \ref{ch:unif} is proved complete with substantial
restrictions on the order and kind of applied rules. To further
remove redundancy, we plan to find alternative, less explosive unification rules
that do not compromise completeness. Furthermore, we want to investigate if
there are more combinations of unification rules that lead to redundant unifiers
and find ways to remove those redundancy-inducing combinations.

We also described only one pragmatic variant of the unification procedure. This
variant was designed after we sifted through hundreds of benchmarks to get a
taste of what unifiers occur in practice. It is possible that this pragmatic variant
can be further tweaked to perform more efficiently without loosing many useful unifiers.

Lastly, we fixed the order in which Zipperposition applies different unification
rules. It would be interesting to see if applying imitation before projection
(or vice versa) yields performance improvements on some kinds of problems. 
We also postponed the application of the most explosive rules such as iteration
to the very end. However, it would not be surprising that there are many hard hand-crafted
problems on which applying these rules early on would lead to a proof more quickly.

\ourpara{Pragmatic Techniques and Heuristics } Many of the Boolean reasoning
techniques described in Chapter \ref{ch:bools} were either inspired by the ones
implemented in traditional higher-order theorem provers or by the unsolved TPTP
benchmarks. Since the work present in Chapter \ref{ch:bools} anticipated the
complete higher-order calculus that could make these techniques obsolete, we
implemented only the most successful ones that are easy to port to
saturation-based provers.  Most notably, unlike Satallax
\cite{cb-2013-satallax}, Zipperposition does not use the fact that there only
finitely many functions of a type built using only the function type constructor
($\rightarrow$) and Boolean type ($o$). Implementing techniques for efficient
enumeration of all Boolean functions of a given type could help Zipperposition
prove problems that are still out of its reach, but can be solved by Satallax.

In Chapter \ref{ch:ho-techniques} we explored how we keep the explosion inherent
to the \osup{} calculus under control. Due to time and resource
constraints, in the evaluation of our new given-clause loop we fixed some
parameters that control which streams are queried. We plan to tune these
parameters in the future.  Similarly, we provided only some of the clause
priority functions that rely on higher-order features. We plan to investigate
more proofs in detail to create new priority functions that more cautiously
separate less and more useful higher-order inferences. Lastly, we observed that
use of Ehoh as a backend greatly improves the performance of Zipperposition. We
plan to further investigate the proofs that Ehoh returns and find out how
Zipperposition can find them without the use of a backend.

\ourpara{Further Extensions of \ehohii{}} \ehohii{} outperforms other provers on
problems coming from proof assistants. However, its performance on TPTP problems is
slightly behind Zipperposition. We plan to further investigate the benchmarks on
which \ehohii{} fails, but that are proved by Zipperposition. One of the first
features we plan implement is dynamic clausification, which can possibly help
improve \ehohii{}'s performance on first-order problems as well. We also plan to
experiment with implementing some of the \osup{} rules necessary for
completeness such as \infname{FluidSup}. Heuristics for \ehohii{} are trained on
a subset of Sledgehammer and TPTP problems. We plan to use other benchmark sets
such as GRUNGE \cite{bgksu-19-grunge} to make \ehohii{} efficient on a wider range of higher-order
benchmarks.

\ourpara{SAT-inspired eliminations for superposition} We already implemented
some of the techniques described in Chapter~\ref{ch:satfol} in E. However, this
implementation is not properly tested and optimized. We plan to put finishing
touches to this implementation and test the effects of implemented preprocessing
techniques on E's performance. There are more techniques used in SAT solving,
notably bounded variable addition \cite{mhb-12-reencoding} and blocked clause
addition \cite{ok-99-er}, that we did not port to first-order logic. We want to
investigate whether these techniques can be lifted to first-order logic. Lastly,
we already started work on porting existing techniques to higher-order logic.
